### 2015-07-10 Data/Datahandling.Rmd
All three RECS years in one file.  Work on the individual columns next.

---
title: "Multiyear RECS data handling"
author: "James Woods"
date: "07/10/2015"
output: html_document
---

```{r}
recs2009_public <- read.csv("~/Research/EnergyEthnicity/Data/2009/recs2009_public.csv")

RECS05alldata <- read.csv("~/Research/EnergyEthnicity/Data/2005/RECS05alldata.csv")

```
Those two are simple but the 2001 data comes in many parts

```{r}
publicuse102001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse102001.csv", header=TRUE)  
publicuse2001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse2001.csv", header=TRUE)   
publicuse42001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse42001.csv", header=TRUE)  
publicuse72001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse72001.csv", header=TRUE)
publicuse112001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse112001.csv", header=TRUE)  
publicuse22001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse22001.csv", header=TRUE)  
publicuse52001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse52001.csv", header=TRUE)  
publicuse82001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse82001.csv", header=TRUE)
publicuse122001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse122001.csv", header=TRUE) 
publicuse32001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse32001.csv", header=TRUE)
publicuse62001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse62001.csv", header=TRUE)  
publicuse92001<-read.csv("~/Research/EnergyEthnicity/Data/2001/publicuse92001.csv", header=TRUE)

#DOEID is the merge key
library(reshape)
All2001<-merge_recurse(list(publicuse102001,publicuse112001,publicuse122001,publicuse2001,publicuse22001,publicuse32001,publicuse42001,publicuse52001,publicuse62001,publicuse72001,publicuse82001,publicuse92001))

```

Harmonizing these into a single dataframe will be hard.  Lets check the names in each file and find the commonalities in the names.

```{r}
intersect(intersect(names(All2001),names(RECS05alldata)),names(recs2009_public))

```
Plenty of common names.  Lets look at the union.

```{r}
union(union(names(All2001),names(RECS05alldata)),names(recs2009_public))

```
So, we have <400 common names out of ~1500 total.

Step one.  Make sure we have a unique ID accross all years for the observations.

```{r}

All2001$ID<-paste(as.character(All2001$DOEID),"2001",sep="")
RECS05alldata$ID<-paste(as.character(RECS05alldata$DOEID),"2005",sep="")
recs2009_public$ID<-paste(as.character(recs2009_public$DOEID),"2009",sep="")
```

And give an indicator of survey year

```{r}
All2001$Year<-2001
RECS05alldata$Year<-2005
recs2009_public$Year<-2009
```

Now lets combine the columns that are common into one dataframe

```{r}
CommonCols<-intersect(intersect(names(All2001),names(RECS05alldata)),names(recs2009_public))
RECS<-rbind(All2001[CommonCols],RECS05alldata[CommonCols],recs2009_public[CommonCols])
```

Check the number of observations.
```{r}
dim(All2001)
dim(RECS05alldata)
dim(recs2009_public)
dim(RECS)
```
They all checkout.

Now lets merge in the columns that are unique to each dataframe.
```{r}
#Need unique names plus ID only in the merge.

outersect <- function(x, y) {
  sort(c(setdiff(x, y),
         setdiff(y, x)))
}

U2001<-c("ID",outersect(names(RECS),names(All2001)))
U2005<-c("ID",outersect(names(RECS),names(RECS05alldata)))
U2009<-c("ID",outersect(names(RECS),names(recs2009_public)))

RECS<-merge(RECS,All2001[U2001],by="ID", all.x=TRUE)
RECS<-merge(RECS,RECS05alldata[U2005],by="ID", all.x=TRUE)
RECS<-merge(RECS,recs2009_public[U2009],by="ID", all.x=TRUE)
```

That should do it.

```{r}
summary(RECS)

```